Predict Disease Outcome by Community Structures

### 2016-08-29 code/HumanCdGF_microbiota_predict.Rmd
reduce features in mor/pers RF class model

# add log transformed CFU
meta_file$log_cfu <- log10(meta_file$cdiff_cfu + 1)
# remove cecum samples
meta_file <- meta_file[! meta_file$sample_type == 'cecum', ]
# remove unused cage P2 and non 431 cdiff
meta_file <- meta_file[! meta_file$cage_id %in% c('P2', 'CON', 'CON2', 'CON3'), ]

@@

##Can we predict moribundity with the community structure at Day 0?
```{r fig.width=10, fig.height=5}
Predict_early_euth_df <- data.frame(Euth_Early = ifelse(meta_file[meta_file$day %in% 0,
                                                                  'Euthanized'] %in% 10,
                                                        'No','Yes'), 
                                    row.names = meta_file[meta_file$day %in% 0, 'Mouse_ID'])
Predict_early_euth_df <- merge(Predict_early_euth_df, rel_abund_stool, by='row.names')
mouse_list <- Predict_early_euth_df$Row.names
Predict_early_euth_df <- Predict_early_euth_df[,!(colnames(Predict_early_euth_df) %in% "Row.names")]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df, 
                              importance = TRUE, ntree = n_trees)

unpredicted <- mouse_list[Predict_early_euth_df$Euth_Early != rf_early_euth$predicted]
rf_euth_import <- data.frame(importance = head(
  sort(rf_early_euth$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 30))
ROC_euth <- roc(as.numeric(Predict_early_euth_df$Euth_Early), as.numeric(rf_early_euth$predicted))
roc_plot <- data.frame(Specificity = 1 - ROC_euth$specificities, Sensitivity = ROC_euth$sensitivities)

euth_rf_plot <- ggplot(data = roc_plot, aes(x = Specificity, y = Sensitivity)) + geom_path() + 
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed', alpha = 0.5)
euth_import_plot <- ggplot(data = rf_euth_import, aes(x = factor(rownames(rf_euth_import)), y = importance)) + 
        geom_point() + scale_x_discrete(limits = rev(rownames(rf_euth_import))) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()
# title = 'OTU Importance in Prediction of Day Euthanized from Day 0 Community',

ggdraw()+
  draw_plot(euth_rf_plot, 0, 0, 0.5, 1) + 
  draw_plot(euth_import_plot, 0.5, 0, 0.5, 1)
```

Confusion Matrix  
`r kable(rf_early_euth$confusion)`

OOB error rate = `r 100*tail(rf_early_euth$err.rate[ , 1], 1)`

Mice which were euthanized early, but the model predicts the will were not (Mouse Tag - Cage):
`r unpredicted`

>>>>

# add log transformed CFU
meta_file$log_cfu <- log10(meta_file$cdiff_cfu + 1)
# remove cecum samples
meta_file <- meta_file[meta_file$sample_type == 'stool', ]
# remove unused cage P2 and 
meta_file <- meta_file[! meta_file$cage_id %in% 'P2', ]
# select only mice with 431 cdiff
meta_file <- meta_file[meta_file$cdiff_strain == '431', ]

@@

##Can we predict moribundity with the community structure at Day 0?
```{r fig.width=10, fig.height=5}
Predict_early_euth_df <- data.frame(Euth_Early = ifelse(meta_file[meta_file$day %in% 0,
                                                                  'Euthanized'] %in% 10,
                                                        'No','Yes'), 
                                    row.names = meta_file[meta_file$day %in% 0, 'Mouse_ID'])
Predict_early_euth_df <- merge(Predict_early_euth_df, rel_abund_stool, by='row.names')
mouse_list <- Predict_early_euth_df$Row.names
Predict_early_euth_df <- Predict_early_euth_df[,!(colnames(Predict_early_euth_df) %in% "Row.names")]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df, 
                              importance = TRUE, ntree = n_trees)
rf_euth_import <- data.frame(importance = head(sort(rf_early_euth$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 10))

rf_early_euth_top_otu <- randomForest(as.factor(Euth_Early) ~ ., 
                                      data = Predict_early_euth_df[, c(rownames(rf_euth_import), 'Euth_Early')], 
                              importance = TRUE, ntree = n_trees)

unpredicted_top_otu <- mouse_list[Predict_early_euth_df$Euth_Early != rf_early_euth_top_otu$predicted]
rf_euth_import_top_otu <- data.frame(importance = sort(rf_early_euth_top_otu$importance[ , 'MeanDecreaseAccuracy'], 
                                                       decreasing = T))
ROC_euth_top_otu <- roc(as.numeric(Predict_early_euth_df$Euth_Early), as.numeric(rf_early_euth_top_otu$predicted))
roc_plot_top_otu <- data.frame(Specificity = 1 - ROC_euth_top_otu$specificities, 
                       Sensitivity = ROC_euth_top_otu$sensitivities)

euth_rf_plot_top_otu <- ggplot(data = roc_plot_top_otu, aes(x = Specificity, y = Sensitivity)) + 
                      geom_path() + geom_abline(intercept = 0, slope = 1, 
                                                linetype = 'dashed', alpha = 0.5)
euth_import_plot_top_otu <- ggplot(data = rf_euth_import_top_otu, 
                           aes(x = factor(rownames(rf_euth_import_top_otu)), y = importance)) + 
        geom_point() + scale_x_discrete(limits = rev(rownames(rf_euth_import_top_otu))) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()
# title = 'OTU Importance in Prediction of Day Euthanized from Day 0 Community',

ggdraw()+
  draw_plot(euth_rf_plot_top_otu, 0, 0, 0.5, 1) + 
  draw_plot(euth_import_plot_top_otu, 0.5, 0, 0.5, 1)
```

Confusion Matrix  
`r kable(rf_early_euth_top_otu$confusion)`

OOB error rate = `r 100*tail(rf_early_euth_top_otu$err.rate[ , 1], 1)`

Mice which were euthanized early, but the model predicts the will were not (Mouse Tag - Cage):
`r unpredicted_top_otu`
